---
title: "Python is cool!"
author: "Kylie Wakefield"
date: "5/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::knit_engines$set(python = reticulate::eng_python)
library(reticulate)
use_python("/Users/kyliewakefield/Documents/UT/Spring2020/Website/python/bin/python")
```

Python and R can interact together by utilizing the reticulate package. Through this package, the {r} and {python} code chunks will talk to one another. 

One cool thing about python is it's ability to create strings. The strings can be a collection of words or phrases. Here are some of my favorite foods...
```{python}
#python code chunk

'ice cream, rasberries, breakfast tacos, and queso'
```

They can also be saved as an object, like R!
```{python}
#python code chunk

favoritefoods = "ice cream, rasberries, breakfast tacos, and queso "
favoritefoods
```

When r and python interact together, the two objects can be shown by using the print() function. The print() function also gives a cleaner output of the strings. (r."object") allows the python code to output the r object allow with the string in the python code. Pretty cool! 
```{r}
#r code chunk

food<-"my favorite foods are"
```

```{python}
#python code chunk

print(r.food,favoritefoods)
```

If I wanted to convert all of my favorite foods into upper case letters, then I would  use the str.upper() function. This is really emphasizing my favorite foods. YUM!!!
```{python}
#python code chunk

uppercase = str.upper(favoritefoods)
uppercase
print(r.food, uppercase)
```

I hope that through my blog you learned a little about python and r studio, as well as my favorite foods. Maybe this while inspire you to indulge in your favorites... :)